---
title: "Lecture 7: Joint Distribution"
subtitle: "STA237: Probability, Statistics and Data Analysis I"
author: Michael J Moon
output: html_document
---


```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
```


## Simulating Joint Distributions

Suppose $R$ and $\theta$ are two continuous random variables with the following distributions.

$$R\sim\text{U}(0, 1)\quad\text{and}\quad\theta\sim\text{U}(0, \pi)$$

Let $X=R\cos(\theta)$ and $Y=R\sin(\theta)$. Simulate 1,000 pairs of $(X,Y)$.


```{r}
# Write your R codes here.
```

## Plotting Simulated Values

The following code plots a semi circle with radius 1. Add the simulated points $(x_i,y_i)$ for $i=1,\dots,1000$ on to the plot.

```{r fig.asp=2/3}
circlei <- seq(0, pi, 0.01)
plot(cos(circlei), sin(circlei), type = "l",
     xlab = "x", ylab = "y")
```


Alternatively, in `ggplot`...

```{r fig.asp=2/3}
ggplot() +
  theme_classic() +
  geom_line(aes(x = cos(circlei), y = sin(circlei))) +
  labs(x = "x", y = "y")
```


## Knitting a RMarkdown Document

To knit the document, click **Knit** button above the document area or pres **Ctr + Shift + K**. If you wish to render a PDF file, change `output: html_document` at the top of the document to `output:pdf_document`. 

