---
title: "Lecture 7: Joint Distribution"
output: 
  learnr::tutorial:
    css: css/styles.css
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
set.seed(237)
```

## R Preparation

### Using a table in R

So far, we have worked with vectors and matrices which can save values of one data type. 
We can create tables consisting of columns in different data types via `data.frame()` in R.

For example, the data frame below has a column of numbers and a column of characters.

```{r echo=TRUE}
some_table <- data.frame(
  numbers = 1:5,
  alphabets = letters[1:5],
  l = sample(c(T, F), 5, replace = TRUE),
  y = rpois(5, 100),
  w = rgeom(5, 1/100)
)
some_table
```

Many functions in R, including `ggplot()`, are designed to work with data frames. For example, in `ggplot()`,
the first argument is expected to be a table^[We have been manually bypassing the first argument with `mapping = aes()`.]
Once we pass in a table, you can define the mappings using the column names. 

```{r echo=TRUE}
library(ggplot2)
ggplot(some_table,          # provide the table
       aes(x = numbers)) +  # map the column `numbers` to x-axis
  theme_bw() +
  geom_point(aes(y = ifelse(l, y, w), # draw points using `y` or `w` based on the value of `l`
                 color = l)) # change color based on the value of `l`
```

> Specifying aesthetics such as `color`, `size`, `linetype`, `shape`, etc. inside `aes()` changes the specified aesthetics based on the mapped values.

Most multivariate data come in a table format and R makes it easy to work with them.

+  You can access elements of a data frame using 2D indexes in the same way you use 2D indexes with matrices.

```{r echo=TRUE}
some_table[1, ]  # extracting the first row
some_table[1:3, ]  # extracting the first three rows
some_table[ , 3] # extracting the third column
some_table[1, 3:4] # extracting the element in the first row, third and fourth columns
```

> When you extract multiple columns, the result is still a data frame. 

+  On the other hand, using a single positional index behaves differently from a matrix and returns a column.

```{r echo=TRUE}
some_table[3]   
```

+  You can also extract a column using `$` followed by the column name.

```{r echo=TRUE}
some_table$l
```


> Using 2D index to extract a column (e.g., `some_table[ , 3]`) returns a vector whereas using 1D index (e.g., `some_table[3]`) or `$` extractor returns a single column data frame. It doesn't affect the codes in most cases but can be a good place to check if you face an unexpected error.


### Practice

In the code chunk below, `some_table` is available for you. 

i. Extract values of `y` from rows where `l` is `TRUE`.
ii. Extract values of `w` from rows where `l` is `FALSE`.
iii. Add the five values extracted.

```{r sometable-setup}
set.seed(237)
some_table <- data.frame(
  numbers = 1:5,
  alphabets = letters[1:5],
  l = sample(c(T, F), 5, replace = TRUE),
  y = rpois(5, 100),
  w = rgeom(5, 1/100)
)
```

```{r sometable, exercise=TRUE, exercise.lines=15}
some_table
```

```{r sometable-hint-1}
# extract the values
some_table$y[some_table$l]
some_table$w[!some_table$l]
```

```{r sometable-solution}
sum(some_table$y[some_table$l]) + sum(some_table$w[!some_table$l])
```

## Simulating a joint distribution

Suppose $R$ and $\theta$ are two continuous random variables with the following distributions.

$$R\sim\text{U}(0, 1)\quad\text{and}\quad\theta\sim\text{U}(0, 2\pi)$$

Let $X=R\cos(\theta)$ and $Y=R\sin(\theta)$. 

i. Simulate 5,000 pairs of $(X,Y)$.
ii. Create a data frame of columns `R`, `theta`, `X`, `Y`.
iii. Plot `X` vs. `Y` as points using the data frame.


```{r jsim, exercise=TRUE, exercise.lines=15}
N <- 5000
```

```{r jsim-hint-1}
N <- 5000
R <- runif(N)
theta <- runif(N, max = 2 * pi) # pi is a predefined constant in R
```

```{r jsim-hint-2}
N <- 5000
R <- runif(N)
theta <- runif(N, max = 2 * pi) # pi is a predefined constant in R
sims <- data.frame(
  R = R, theta = theta,
  X = R * cos(theta), Y = R * sin(theta)
)
```


```{r jsim-hint-3}
N <- 5000
R <- runif(N)
theta <- runif(N, max = 2 * pi) # pi is a predefined constant in R
sims <- data.frame(
  R = R, theta = theta,
  X = R * cos(theta), Y = R * sin(theta)
)
ggplot(sims, aes(x = X, y = Y)) +
  theme_classic() +
  geom_point()
```

```{r jsim-hint-4}
N <- 5000
R <- runif(N)
theta <- runif(N, max = 2 * pi) # pi is a predefined constant in R
sims <- data.frame(
  R = R, theta = theta,
  X = R * cos(theta), Y = R * sin(theta)
)
ggplot(sims, aes(x = X, y = Y)) +
  theme_void() +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  coord_equal() + # plot x and y axis at equal scales
  geom_point(alpha = 0.5) # semi-transparent points
```

> Optionally, set `alpha = 0.5` inside `geome_point()` to make the points semi-transparent.
